---
title: "Halloween Candy Data"
output: html_notebook
---

```{r}
# loading in the libraries
library(tidyverse)
library(dplyr)
library(here)
library(stringr)
```


```{r}
# loading in the data, preserving the column types frome the cleaning script
here::here()
candy_clean <- read_csv(here("clean_data/candy_clean.csv"), col_types = list(
  candy = col_character(), 
  rating = col_factor(), 
  year = col_integer(), 
  country = col_character(), 
  age = col_double(), 
  gender = col_factor(), 
  gender = col_factor(), 
  trick_or_treating = col_character()
  ))
```

1.  What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
```{r}
q1_number_of_ratings <- candy_clean %>% 
  select(rating) %>% 
  filter(!is.na(rating)) %>% 
  summarise(num_of_candy_ratings = n()) %>% 
  pull()

paste("The total number of candy ratings given across the three years is",formatC(q1_number_of_ratings,big.mark = ","))
# first answer is 742,764
```


2.  What was the average age of people who are going out trick or treating?
```{r}
q2_trick_or_treating <- candy_clean %>% 
  select(age, trick_or_treating) %>% 
  filter(
    !is.na(trick_or_treating),
    trick_or_treating == "Yes"
    ) %>% 
  summarise(average_age = mean(age, na.ram = TRUE)) %>% 
  pull()

paste("The average age of people who are going out trick or treating is", round(q2_trick_or_treating, 0), "years old.")
# first answer is 35 years old
```


3.  What was the average age of people who are not going trick or treating?
```{r}
q3_trick_or_treating <- candy_clean %>% 
  select(age, trick_or_treating) %>% 
  filter(
    !is.na(trick_or_treating),
    trick_or_treating == "No"
    ) %>% 
  summarise(average_age = mean(age, na.ram = TRUE)) %>% 
  pull()

paste("The average age of people who are not going out trick or treating is", round(q3_trick_or_treating, 0), "years old.")
# first answer is 39 years old
```

4. How many people rated Starburst as despair?
```{r}
q4_starburst_as_despair <- candy_clean %>% 
  select(candy, rating) %>% 
  filter(
    !is.na(rating),
    candy == "starburst",
    rating == "DESPAIR"
    ) %>% 
  summarise(count = n()) %>% 
  pull()

paste(formatC(q4_starburst_as_despair,big.mark = ","), "people rated Starburst as despair.")
# first answer is 1,872 people
```

For the next three questions, count despair as -1, joy as +1 and meh as 0.
```{r}
glimpse(candy_clean)
q5_ratings_recoded <- candy_clean %>% 
  mutate(
    rating = recode_factor(rating, JOY = 1, MEH = 0, DESPAIR = -1),
    as.numeric(rating)
  ) %>% 
   filter(
    !is.na(rating))
class(q5_ratings_recoded$rating)
```


5. What was the most popular candy bar by this rating system for each gender in the dataset?
```{r}
q5_answer <- q5_ratings_recoded %>% 
  select(candy, rating, gender) %>% 
  filter(gender == "Male") %>% 
  group_by(candy) %>% 
  summarise(popularity = sum(rating)) %>% 
  slice_max(n = 1)
```


6. What was the most popular candy bar in each year?
7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
